ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-07 rebuttal.tex

\subsection*{Abstract and introduction}

\begin{reviewer}
    %
    4. In the abstract, it is mentioned that the Python-based package PyKED creates ChemKED-
    formatted files. However in the article there is no evidence that PyKED aides in creating
    ChemKED-formatted files.
    %
\end{reviewer}

\begin{response}
    %
    The reviewer is correct, and this was a typo in the abstract. We have clarified that PyKED is
    intended as software to validate and work with ChemKED files.

    That said, we do have ongoing work to create a graphical user interface to create ChemKED files,
    as this was requested by several of the experimental teams that have expressed interest in the
    ChemKED format. The work is available publicly at
    \url{https://github.com/pr-omethe-us/ChemKED-gui}. In addition, we have a built-in converter
    for generating ChemKED files from existing ReSpecTh files.
    %
\end{response}

@@

\begin{reviewer}
    %
    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
    reference to it on the GitHub wiki roadmap of PyKED.
    %
\end{reviewer}

\begin{response}
    %
    Thank you for the comment, we have corrected this oversight in the roadmap and the
    conclusions (\textbf{page 17}).
    %
\end{response}

### 2017-10-07 rebuttal.tex

\begin{reviewer}
    %
    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
    reference to it on the GitHub wiki roadmap of PyKED.
    %
\end{reviewer}

\begin{response}
    %
    Thank you for the comment, we have corrected this oversight in the roadmap and the
    conclusions (\textbf{page 17}).
    %
\end{response}

\begin{reviewer}
    %
    6. I don't completely agree with the second part of the statement on page 2 line 16-18:
    ``Second, the XML format is intended to be a machine-readable mark-up language rather than a
    data format, and its lack of human readability presents a barrier to creating and working with
    database files.''

    In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
    however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
    read. If XML would lack human readability as it is stated in the article, it would not be such a
    popular format, and I don't think PrIMe and ReSpecTh would be using it.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that XML is technically human readable by an expert (as opposed to,
    say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
    mean human-\emph{understandable}. We have modified the text on \textbf{page 2} to clarify this
    point; please see also the response to a similar comment from Reviewer \#1.
    %
\end{response}

### 2017-10-07 rebuttal.tex

    In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
    however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
    read. If XML would lack human readability as it is stated in the article, it would not be such a
    popular format, and I don't think PrIMe and ReSpecTh would be using it.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that XML is technically human readable by an expert (as opposed to,
    say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
    mean human-\emph{understandable}. We have modified the text on \textbf{page 2} to clarify this
    point; please see also the response to a similar comment from Reviewer \#1.
    %
\end{response}

\begin{reviewer}
    %
    7.~An initial version of the ChemKED format was introduced by Niemeyer, as is mentioned in the
    introduction. The mentioned article describes a Python-package called PyTeCK. How does this
    PyTeCK package relate to the PyKED package that is presented in the submitted article?
    %
\end{reviewer}

\begin{response}
    %
    PyTeCK is a software package for automatically validating chemical kinetic models based
    on experimental data encoded in the ChemKED format; thus, PyTeCK relies on the PyKED software
    to interpret and validate ChemKED files.
    %
\end{response}

@@

\begin{reviewer}
    %
    8.~Page 3, line 10: ``Our motivations are similar to those of '', `to' is missing.
    %
\end{reviewer}

\begin{response}
    %
    We have edited that sentence on \textbf{page 3} to correct the error.
    %
\end{response}

### 2017-10-07 rebuttal.tex

\subsection*{Overview of ChemKED format}

\begin{reviewer}
    %
    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
    to make the community to adopt the ChemKED file structure, it might be easier to do so if
    experimentalists can use the structure internally as well, i.e. to read, store, and share
    non-published internal data.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field. We have
    modified the ChemKED schema in response to this comment so that the only required subfields of
    the reference field are the authors and the year.

    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
    This option turns off all validation, including the validation of the reference material
    and enforcement of required fields. We recognize that this is a limitation at present, if users
    wish to, for example, validate that their units are correct but do not wish to validate some
    other aspect of the file. Work is ongoing to improve this situation, and discussion is contained
    in a public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
    %
\end{response}

@@

\begin{reviewer}
    %
    10.~Page 4, line 32: ``The second section of the file encodes the experimental data''. Perhaps
    it is easier to follow if this is briefly mentioned before the authors explain the details of
    the first section of the ChemKED format.
    %
\end{reviewer}

\begin{response}
    %
    We have updated the text on \textbf{page 4} to incorporate the suggested change.
    %
\end{response}

### 2017-10-07 pyked.tex
response to close #28

The second section of the file encodes the experimental data as a sequence of mappings. The
top-level key for the sequence is called \yabox|datapoints|. For the current version of the \ck{}
schema, only autoignition delay experiments are supported. The following information os required in
each element of the sequence:
%
\begin{itemize}
    \item \yabox|temperature| (required, sequence): The temperature of the
    experiment, with units and optionally uncertainty
    \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
    experiment, with units and optionally uncertainty
    \item \yabox|pressure| (required, sequence): The pressure of the experiment,
    with units and optionally uncertainty
    \item \yabox|first-stage-ignition-delay| (optional, sequence): If two stages of ignition are
    present in the experiment, this is the length of the first stage of ignition, and the
    \yabox|ignition-delay| is then the overall ignition delay
    \item \yabox|equivalence-ratio| (optional, float): The value of the equivalence ratio
    \item \yabox|composition| (required, mapping): The composition of the
    mixture in the experiment, described via a nested mapping with the fields:
    \begin{itemize}
        \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
        \item \yabox|species|: sequence of mappings, with the fields:
        \begin{itemize}
            \item \yabox|species-name| (required, string): The name of the species
            \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
            \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
            \yabox|elemental-composition| (required, string or sequence): The InChI
            or SMILES string representing the molecule, or its atomic or elemental
            composition (\yabox|atomic-composition| and\\
            \yabox|elemental-composition| are synonyms) as a sequence
        \end{itemize}
    \end{itemize}
    \item \yabox|ignition-type| (required, mapping): The method used to detect
    ignition in the experiments. The required fields are
    \begin{itemize}
        \item \yabox|type| (required, string): How ignition delay was
        measured; one of \yabox|d/dt max| (indicates the ignition point was
        found at the maximum of the time derivative of the \yabox|target|),
        \yabox|max| or \yabox|min| (indicates the ignition point was at the
        maximum or minimum of the \yabox|target|), \yabox|1/2 max| (
        indicates the half-maximum point of \yabox|target|), or \yabox|d/dt max extrapolated|
        (indicates the maximum slope of the target extrapolated to the baseline)
        \item \yabox|target| (required, string): The target for the ignition
        \yabox|type| measurement; one of \yabox|temperature|, \yabox|pressure|,
        \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
    \end{itemize}
\end{itemize}

>>>>

The second section of the file encodes the experimental data as a sequence of mappings. The
top-level key for the sequence is called \yabox|datapoints|. For the current version of the \ck{}
schema, only autoignition delay experiments are supported. The following information os required in
each element of the sequence:
%
\begin{itemize}
    \item \yabox|temperature| (required, sequence): The temperature of the
    experiment, with units and optionally uncertainty
    \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
    experiment, with units and optionally uncertainty
    \item \yabox|pressure| (required, sequence): The pressure of the experiment,
    with units and optionally uncertainty
    \item \yabox|first-stage-ignition-delay| (optional, sequence): If two stages of ignition are
    present in the experiment, this is the length of the first stage of ignition, and the
    \yabox|ignition-delay| is then the overall ignition delay
    \item \yabox|equivalence-ratio| (optional, float): The value of the equivalence ratio
    \item \yabox|composition| (required, mapping): The composition of the
    mixture in the experiment, described via a nested mapping with the fields:
    \begin{itemize}
        \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
        \item \yabox|species|: sequence of mappings, with the fields:
        \begin{itemize}
            \item \yabox|species-name| (required, string): The name of the species
            \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
            \item \yabox|InChI|, \yabox|SMILES|, or \yabox|atomic-composition|
            (required, string or sequence): The InChI
            or SMILES string representing the molecule, or its atomic
            composition as a sequence
        \end{itemize}
    \end{itemize}
    \item \yabox|ignition-type| (required, mapping): The method used to detect
    ignition in the experiments. The required fields are
    \begin{itemize}
        \item \yabox|type| (required, string): How ignition delay was
        measured; one of \yabox|d/dt max| (indicates the ignition point was
        found at the maximum of the time derivative of the \yabox|target|),
        \yabox|max| or \yabox|min| (indicates the ignition point was at the
        maximum or minimum of the \yabox|target|), \yabox|1/2 max| (
        indicates the half-maximum point of \yabox|target|), or \yabox|d/dt max extrapolated|
        (indicates the maximum slope of the target extrapolated to the baseline)
        \item \yabox|target| (required, string): The target for the ignition
        \yabox|type| measurement; one of \yabox|temperature|, \yabox|pressure|,
        \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
    \end{itemize}
\end{itemize}
